---
title: "MP Oyster Depuration Project"
author: "Garth C"
date: "December 6, 2018"
output: html_document
---

Load the libraries we'll need

```{r}
library(plyr)
library(reshape2)
library(ggplot2)
library(glmmADMB)
library(MuMIn)
```

Import the oyster depuration data and view it.

```{r}
oysterdep <- read.csv("~/GRAD school/ACRDP/Depuration/Data/Analysis/R code/oysterdep.csv")

head(oysterdep)
```

Get all of the variables into the right format.

```{r}
oysterdep$id <- as.factor(oysterdep$id)
oysterdep$sampleday <- as.factor(oysterdep$sampleday)
oysterdep$size.cat <- as.factor(oysterdep$size.cat)
oysterdep$Table <- as.factor(oysterdep$Table)
oysterdep$Tank <- as.factor(oysterdep$Tank)
oysterdep$observer <- as.factor(oysterdep$observer)
```

Make a new dataframe to work with and rename size categories.

```{r}
oysterdep2 <- oysterdep

oysterdep2$size.cat <- factor(oysterdep2$size.cat, 
                              levels=c('twentyto50', 'fiftyto100',
                                       'onehundredto500', 'fivehundredto1000', 
                                       'onethousandto5000'))  # reorder


oysterdep2$size.cat <- mapvalues(oysterdep2$size.cat, 
                                 from = c('twentyto50', 'fiftyto100', 
                                          'onehundredto500', 
                                          'fivehundredto1000', 
                                          'onethousandto5000'), 
                                 to = c('20-50', '50-100', '100-500',
                                        '500-1000', '1000-5000'))  # rename
summary(oysterdep2$size.cat)  # view
```

We need to melt the dataframe so that the particle types are organized by row instead of by column.

```{r}
oysterdep3 <- melt(oysterdep2,
                   id.vars = c('id', 'sampleday', 'size.cat', 'Table', 'Tank',
                               'species', 'DOFSample', 'length', 'width', 
                               'depth', 'cont.weight', 'cont.dryweight', 
                               'dry.weight', 'shell.dry.weight', 'CI',
                               'time.in.oven', 'time.out.oven',
                               'dat.filter', 'dat.count', 'observer'))

##  Take out unnecessary columns

oysterdep3 <- oysterdep3[c('id', 'sampleday', 'size.cat', 'Table', 'Tank', 
                         'DOFSample', 'length', 'width', 'depth', 'dry.weight',
                         'shell.dry.weight', 'CI', 'dat.filter', 'dat.count',
                         'observer', 'variable', 'value')]

head(oysterdep3)  # now 'variable' is particle category and 'value is the count'
```

Now we want to see what kind of particles we found and remove any categories that we didn't end up using.

```{r}
cat.sums <- aggregate(oysterdep3$value ~ 
                        oysterdep3$variable, FUN = sum)  # sums everything
print(cat.sums)
```

So we can remove the purple fibre, gray fibre, yellow fragment, purple fragment, gray fragment, brown fragment, orange fragment, and black fragment categories because we didn't count any of these.

```{r}
oysterdep4 <- subset(oysterdep3, variable != 'purp.fib' & 
                       variable != 'gray.fib' & variable != 'yell.frag' &
                       variable != 'purp.frag' & variable != 'gray.frag' & 
                       variable != 'brown.frag' & variable != 'orang.frab' &
                       variable != 'black.frag')  # removes all these rows

oysterdep4$variable <- as.character(oysterdep4$variable)
oysterdep4$variable <- as.factor(oysterdep4$variable)  # gets rid of those names

summary(oysterdep4$variable)  # view
```

Let's make the category names a bit more useful for plotting

```{r}
oysterdep4$variable <- mapvalues(oysterdep4$variable,
                                 from = c(levels(oysterdep4$variable)), 
                                 to = c('Black fibres', 'Blue fibres', 
                                 'Blue fragments', 'Brown fibres', 
                                 'Clear fibres', 'Clear fragments', 
                                 'Clear spherules', 'Green fibres',
                                 'Green fragments', 'Orange fibres',
                                 'Orange fragments', 'Pink fibres', 
                                 'Pink fragments', 'Red fibres',
                                 'Red fragments', 'Turquoise fibres', 
                                 'Turquoise fragments', 'White spherules', 
                                 'Yellow fibres'))

## And reorder them
oysterdep4$variable <- factor(oysterdep4$variable, 
                              levels = c('Black fibres', 'Blue fibres',
                                         'Brown fibres', 'Clear fibres',
                                         'Green fibres', 'Orange fibres',
                                         'Pink fibres', 'Red fibres',
                                         'Turquoise fibres',
                                         'Yellow fibres', 'Blue fragments',
                                         'Clear fragments', 'Green fragments',
                                         'Orange fragments', 'Pink fragments',
                                         'Red fragments', 'Turquoise fragments',
                                         'Clear spherules', 'White spherules'))

summary(oysterdep4$variable)
```

Take total sums and plot the results.

```{r}
totalsums <- ddply(oysterdep4, c('id', 'sampleday', 'Table', 'Tank', 
                                 'length', 'width', 'depth', 'dry.weight',
                                 'shell.dry.weight', 'CI', 'observer'),
                   summarise, sum = sum(value))  # adds all the counts together

## Plot according to count per individual

ggplot(totalsums, aes(x=sampleday , y=sum)) + 
  geom_boxplot(position=position_dodge(), size=1) + 
  xlab('Day') + 
  coord_cartesian(ylim = c(0, 41)) +
  ylab('# APs') +
  guides(fill=FALSE) +
  theme_bw() +
  theme(text = element_text(size=12), 
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

## Plot according to concentration (particles/dry tissue weight)

ggplot(totalsums, aes(x=sampleday , y=sum/dry.weight)) + 
  geom_boxplot(position=position_dodge(), size=1) + 
  xlab('Day') + 
  coord_cartesian(ylim = c(0, 20)) +
  ylab('# APs/g dry tissue') +
  guides(fill=FALSE) +
  theme_bw() +
  theme(text = element_text(size=12), 
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```

Now load in the blanks dataset and set it up for plotting, same as with the oyster data set. First we load in the data.

```{r}
depblanks <- 
  read.csv("~/GRAD school/ACRDP/Depuration/Data/Analysis/depblanks.csv")

head(depblanks)

depblanks$id <- as.factor(depblanks$id)
depblanks$size.cat <- as.factor(depblanks$size.cat)
depblanks$observer <- as.factor(depblanks$observer)
```

Now we make a new dataframe to work with and rename size categories.

```{r}
depblanks2 <- depblanks

depblanks2$size.cat <- factor(depblanks2$size.cat, 
                              levels=c('twentyto50', 'fiftyto100',
                                       'onehundredto500', 'fivehundredto1000', 
                                       'onethousandto5000'))  # reorder


depblanks2$size.cat <- mapvalues(depblanks2$size.cat, 
                                 from = c('twentyto50', 'fiftyto100', 
                                          'onehundredto500', 
                                          'fivehundredto1000', 
                                          'onethousandto5000'), 
                                 to = c('20-50', '50-100', '100-500',
                                        '500-1000', '1000-5000'))  # rename
summary(depblanks2$size.cat)  # view
```

Again, we melt the dataframe so that the particle types are organized by row instead of by column.

```{r}
depblanks3 <- melt(depblanks2,
                   id.vars = c('id', 'size.cat', 'DOFSample', 'time.in.oven', 
                               'time.out.oven', 'dat.filter', 'dat.count', 
                               'observer'))

##  Take out unnecessary columns

depblanks3 <- depblanks3[c('id', 'size.cat', 'DOFSample', 'dat.filter', 'dat.count',
                         'observer', 'variable', 'value')]

head(depblanks3)  # now 'variable' is particle category and 'value is the count'
```

Now sum by totals across all samples and particles sizes to get an idea of what we're working with.

```{r}
cat.sums.blanks <- aggregate(depblanks3$value ~ 
                               depblanks3$variable, FUN = sum)  # sums everything
print(cat.sums.blanks)
```

So we can remove the red, yellow, green turquoise, purple, gray, brown, orange, and black fibre categories, and all other categories other than blue fragements.

```{r}
depblanks4 <- subset(depblanks3, c(variable == 'blu.fib' | 
                       variable == 'clear.fib' | variable == 'pink.fib' |
                       variable == 'blu.frag'))  # keep only these rows

depblanks4$variable <- as.character(depblanks4$variable)
depblanks4$variable <- as.factor(depblanks4$variable)  # gets rid of those names

summary(depblanks4$variable)  # view
```

Let's make the category names a bit more useful for plotting

```{r}
depblanks4$variable <- mapvalues(depblanks4$variable,
                                 from = c(levels(depblanks4$variable)), 
                                 to = c('Blue fibres', 'Blue fragments', 
                                        'Clear fibres', 'Pink fibres'))

## And reorder them
depblanks4$variable <- factor(depblanks4$variable, 
                              levels = c('Blue fibres', 'Clear fibres',
                                         'Pink fibres', 'Blue fragments'))

summary(depblanks4$variable)
```

Take total sums and plot the results.

```{r}
totalblanksums <- ddply(depblanks4, c('id', 'observer', 'dat.filter', 'DOFSample'),
                   summarise, sum = sum(value))  # adds all the counts together

## Plot according to count per filtration day

ggplot(totalblanksums, aes(x=DOFSample, y=sum)) + 
  geom_boxplot(position=position_dodge(), size=1) + 
  xlab('Day') + 
  ylab('# APs') +
  guides(fill=FALSE) +
  theme_bw() +
  theme(text = element_text(size=12), 
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```

Now load in the water sample data.

```{r}
depwater <- 
  read.csv("~/GRAD school/ACRDP/Depuration/Data/Analysis/depwater.csv")

head(depwater)

depwater$id <- as.factor(depwater$id)
depwater$day <- as.factor(depwater$day)
depwater$size.cat <- as.factor(depwater$size.cat)
depwater$Table <- as.factor(depwater$Table)
depwater$Tank <- as.factor(depwater$Tank)
depwater$observer <- as.factor(depwater$observer)
```

Now we make a new dataframe to work with and rename size categories.

```{r}
depwater2 <- depwater

depwater2$size.cat <- factor(depwater2$size.cat, 
                              levels=c('twentyto50', 'fiftyto100',
                                       'onehundredto500', 'fivehundredto1000', 
                                       'onethousandto5000'))  # reorder


depwater2$size.cat <- mapvalues(depwater2$size.cat, 
                                 from = c('twentyto50', 'fiftyto100', 
                                          'onehundredto500', 
                                          'fivehundredto1000', 
                                          'onethousandto5000'), 
                                 to = c('20-50', '50-100', '100-500',
                                        '500-1000', '1000-5000'))  # rename
summary(depwater2$size.cat)  # view
```

Again, we melt the dataframe so that the particle types are organized by row instead of by column.

```{r}
depwater3 <- melt(depwater2,
                   id.vars = c('id', 'day', 'size.cat', 'Table', 'Tank', 
                               'DOFSample', 'time.in.oven', 'time.out.oven', 
                               'dat.filter',  'dat.count', 'observer'))

##  Take out unnecessary columns

depwater3 <- depwater3[c('id', 'day', 'size.cat', 'Table', 'Tank', 'DOFSample', 
                         'dat.filter', 'dat.count', 'observer', 'variable', 
                         'value')]

head(depwater3)  # now 'variable' is particle category and 'value is the count'
```

Now sum by totals across all samples and particles sizes to get an idea of what we're working with.

```{r}
cat.sums.blanks <- aggregate(depwater3$value ~ 
                               depwater3$variable, FUN = sum)  # sums everything
print(cat.sums.blanks)
```

So we can remove the yellow, green, purple, gray, pink, brown, orange, and black fibre categories, and all other categories as well.

```{r}
depwater4 <- subset(depwater3, c(variable == 'red.fib' | 
                       variable == 'turq.fib' | variable == 'blu.fib' |
                       variable == 'clear.fib'))  # keep only these rows

depwater4$variable <- as.character(depwater4$variable)
depwater4$variable <- as.factor(depwater4$variable)  # gets rid of those names

summary(depwater4$variable)  # view
```

Let's make the category names a bit more useful for plotting

```{r}
depwater4$variable <- mapvalues(depwater4$variable,
                                 from = c(levels(depwater4$variable)), 
                                 to = c('Blue fibres', 'Clear fibres', 
                                        'Red fibres', 'Turquoise fibres'))

## And reorder them
depwater4$variable <- factor(depwater4$variable, 
                              levels = c('Blue fibres', 'Clear fibres',
                                         'Red fibres', 'Turquoise fibres'))

summary(depwater4$variable)
```

Take total sums and plot the results.

```{r}
totalwatersums <- ddply(depwater4, c('id', 'day', 'Table', 'Tank', 'observer', 
                                     'dat.filter'),
                   summarise, sum = sum(value))  # adds all the counts together

## Plot according to count per filtration day

ggplot(totalwatersums, aes(x=day, y=sum)) + 
  geom_boxplot(position=position_dodge(), size=1) + 
  xlab('Day') +
  ylab('# APs') +
  guides(fill=FALSE) +
  theme_bw() +
  theme(text = element_text(size=12), 
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```

Now fit a model to the oyster data to determine differences over time

```{r}
M1 <- glmmadmb(sum ~ sampleday + (sampleday | Table/Tank), family = 'poisson', 
               data = totalsums) ## Fits a GLMM

summary(M1) # model output
plot(resid(M1) ~ fitted(M1)) # plot residuals vs. fitted
plot(resid(M1) ~ totalsums$sampleday) # plot residuals vs. predictor
qqnorm(resid(M1))
abline(0,1)
```

Looking at the plot of residuals vs. fitted and residuals vs. sampe day, we can see that the variance of the residuals varies according to the AP count and the sampling day. This violates the assumption of homogeneity of variance.

Let's try fitting the same model but assuming a negative binomial distribution.

```{r}
M2 <- glmmadmb(sum ~ sampleday + (sampleday | Table/Tank), family = 'nbinom', 
               data = totalsums) ## Fits a NB GLMM

summary(M2) # model output
plot(resid(M2) ~ fitted(M2)) # plot residuals vs. fitted
plot(resid(M2) ~ totalsums$sampleday) # plot residuals vs. predictor
qqnorm(resid(M2))
abline(0,1)
```

This model is a much better fit, and the variance is much more evenly distributed across the model predictions and across the predictor variable (sampling day).

Now let's try pulling out the predictor (sampling day) and using the anova() function to determine whether sampling day is a significant factor in the model.

```{r}
M3 <- glmmadmb(sum ~ (sampleday | Table/Tank), family = 'nbinom', 
               data = totalsums) ## Same structure except no predictor

anova(M2, M3) # compare the null model with the previously fitted model
```

Sampling day is significant with a p-value of <0.001. We can now go into the model to determine which days are significantly different from each other. Looking back at the summar of M2 we can see that day 3 is the only time where the AP concentrations are not significantly different from day 0 (the reference level).

Now we reorder the levels in sampleday to vary the reference level and figure out which days are significantly different from which other days.

```{r}
totalsums$sampleday <- relevel(totalsums$sampleday, '1')
M2 <- glmmadmb(sum ~ sampleday + (sampleday | Table/Tank), family = 'nbinom', 
               data = totalsums) ## Fits a NB GLMM

summary(M2) # model output
```

Similarly, day 3 is the only day not significantly different from day 1.

```{r}
totalsums$sampleday <- relevel(totalsums$sampleday, '3')
M2 <- glmmadmb(sum ~ sampleday + (sampleday | Table/Tank), family = 'nbinom', 
               data = totalsums) ## Fits a NB GLMM

summary(M2) # model output
```

Only days 5 and 10 are significantly different from day 3.

```{r}
totalsums$sampleday <- relevel(totalsums$sampleday, '5')
M2 <- glmmadmb(sum ~ sampleday + (sampleday | Table/Tank), family = 'nbinom', 
               data = totalsums) ## Fits a NB GLMM

summary(M2) # model output
```

Day 10 is the only day not significantly different from day 5.

So in summary:

Day | 0 | 1 | 3 | 5 | 10
0   | x |sig| - |sig| sig 
1   | x | x | - |sig| sig
3   | x | x | x |sig| sig
5   | x | x | x | x | -
10  | x | x | x | x | x

Now let's replot the data with asterisks used to indicate which days are significantly different from one another.

```{r}
ggplot(totalsums, aes(x=sampleday , y=sum)) + 
  geom_boxplot(position=position_dodge(), size=1) + 
  xlab('Day') + 
  coord_cartesian(ylim = c(0, 41)) +
  ylab('# APs') +
  guides(fill=FALSE) +
  theme_bw() +
  theme(text = element_text(size=12), 
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  annotate("text", x=2, y=5, label = "b")

```